\vspace{-15mm} \fontsize{24pt}{10pt}\selectfont \textbf{Traversing symmetric NAT with predictable port allocation function}

### 2013-12-31 paper_lncs_vashek.tex
finishing paper, introduction

\urldef{\mailsa}\path|xklinec@fi.muni.cz|
\urldef{\mailsb}\path|matyas@fi.muni.cz|
\author{Du\v{s}an Klinec \and Vashek Matyáš}
\institute{Faculty of Informatics, Masaryk University, Brno, Czech Republic,\\
\email{\{xklinec,matyas\}@fi.muni.cz}}

@@

\begin{abstract}
\noindent Network Address Translators cause troubles for VoIP and other P2P services since central servers are needed for communication.
A presence of such potentially malicious hosts in a communication path is not desired, mainly due to increased cost, poor link quality 
and security consequences.
Several solution exist for traversing NAT but symmetric case is still problematic. We propose algorithms using single source port for 
symmetric NAT traversal, each with different properties and applicability. 
Paper gives a theoretic point of view on the NAT traversal problem by stating problem formally and modeling a network connected to 
NAT as a random process, in order to evaluate algorithms in a different settings.
% 
% ToDo; Will be written as the last part of the paper... Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur magna lorem, tempor sed facilisis vel, porta et turpis. Sed et felis a massa dictum posuere. Aliquam hendrerit rhoncus ipsum sit amet placerat. Duis fringilla est eu arcu mollis faucibus non sit amet eros. Vestibulum risus nibh, dapibus vitae laoreet eget, fringilla quis nisl. Proin consequat nibh sit amet mauris suscipit tincidunt. Sed rutrum, purus nec aliquam faucibus, quam libero venenatis nisi, ut tempor mi sapien vel diam. Pellentesque sagittis elit non risus malesuada accumsan. Morbi consequat urna.
\end{abstract}

@@

\section{Introduction}
Network Address Translators (abbreviated as NATs) has been increasing in popularity in recent years, 
mainly as a mean of slowing down IPv4 address space depletion. On one hand, NAT offers
the ability to share a public, globally routable IP address among many internal network hosts. On 
the other hand, it breaks the basic end-to-end principle of the Internet and thus favors server-client 
communication model for clients connected behind NAT.

In many applications is a peer-to-peer (P2P) connection of crucial importance and the quality of a service depends on 
its parameters heavily. Main motivation for this work is Voice-over-IP (VoIP) where direct connection 
between communicating parties is necessary. If P2P connection is not possible, relay 
servers has to be used (TURN~\citep{rfc5766} protocol is used). The presence 
of another node in the communication path can have negative effects on a communication channel quality 
(increased latency, jitter, lower bandwidth), relay server is a potential point of failure, 
has to scale with increasing demands of the network, etc.
Thus relay servers require non-negligible resources what makes them expensive for VoIP providers and
establishing a direct connection between parties is of high importance.

From an attacker perspective, relay servers can be seen as potential enemies trying to sniff and log all 
communication. NAT (especially symmetric one, used as a carrier-grade border firewall in large networks)
forces communicating parties to use such central servers
for communication, making monitoring easier for an attacker. Traversal algorithms has to collaborate 
with central servers in such a way they create a direct link, not using servers any further. 
% 
% some way. They use it to create a direct link, not using servers any further.

In this paper we are focused on a symmetric NAT with a predictable port allocation function. We propose 
new algorithms for establishing direct connection between hosts behind such NATs that are effective and 
has low system requirements.

@@

Some devices with NAT functionality provide special interface enabling traversal, for example NAT Port Mapping Protocol~\citep{rfc6886}. But in general 
it is not widely deployed and it cannot be relied upon in a general case.

ICE\ignore{~\citep{rfc5245}} is used as a technique to traverse NAT in most applications nowadays, 
however does not work with symmetric NATs. ICE mainly uses STUN~\citep{rfc5389} and TURN\ignore{~\citep{rfc5766}} for its operation.

% \section{Related work}
Some interesting algorithms for traversing symmetric NAT were proposed by Y.~Takeda in \citep{takeda}, the case where both clients are behind symmetric NAT
with port sensitive allocation function is not covered. It mentions that sending more packets does not increase a probability of establishing a new connection.
Our motivation is to show that it is actually possible.

RFC~5128 \citep{rfc5128} maps current state of the art in NAT traversal, but does not cover our case. 
The research by Wei~\emph{et.~al.} in \citep{wei}., is devoted to symmetric NAT traversal, but their algorithm is using another technique compared
to our approach. The downside of their algorithm is it requires Time To Live~(TTL) modification of the IP~packet. This is usually considered
as a privileged operation in the operating system and super-user privileges are required for this modification.

Wang~\emph{et.~al.} \citep{Wang:2006:RSN:1156422.1156550} propose a NAT traversal algorithm for scenario where two hosts are
behind symmetric NATs, based on a changing of a source port, while in our approach we want to explore possibilities with constant source port.

In our work we study the situation where two parties -- A,B -- try to establish a direct connection, both of them are behind 
a symmetric NAT (possibly cascade of such NATs) with port sensitive, predictable allocation function. Other scenarios with
predictable allocation function are quite easy to solve and covered in previous research.

\section{Network Address Translation}
The notation used for the rest of the paper follows.

@@

\begin{mydef}
% \begin{align*}
$\mathbb{SC} = \mathbb{IP} \times \mathbb{PORT}$
% \end{align*}%
 is a Socket. $\mathbb{IP}$ is set of all possible IP addresses and $\mathbb{PORT}$ 
is set of all possible ports, basically $\mathbb{PORT} = \{1,2,\dots,65535\}$.
% \begin{align*}
$\mathbb{SP} = \mathbb{IP} \times \mathbb{PORT} \times \mathbb{IP} \times \mathbb{PORT}$
% \end{align*}%
 is a Socket Pair.
\end{mydef}
% 
% \begin{mydef}
% \begin{align*}
% \mathbb{AT} \subseteq \mathbb{SP} \times \mathbb{PORT}
% \end{align*} is an allocation table of the NAT. Second component is external port allocated for particular socket pair.
% \end{mydef}
% Allocation table is a set of all socket pairs and corresponding allocated external port.

\paragraph{NAT and communication.}
NAT connects two networks, denoted as an internal and an external, connected to an internal and an external NAT interface respectively.
Hosts has addresses $\text{IP\textsubscript{in}} \in \mathbb{IP}_{in} \subseteq \mathbb{IP}$ 
and $\text{IP\textsubscript{ex}} \in \mathbb{IP}_{ex} \subseteq \mathbb{IP}$ in the internal and the external network
respectively, $\mathbb{IP}_{in} \cap \mathbb{IP}_{ex} = \emptyset$.
If it is not said otherwise the external NAT address\footnote{We assume NAT has only one external address for simplicity.} is denoted as 
$\text{IP\textsubscript{nat}} \in \mathbb{IP}_{ex}$.

Packet received on the external interface has socket pair {(IP\textsubscript{ex},PORT\textsubscript{ex},IP\textsubscript{nat},PORT\textsubscript{nat})}.
Packet received on the internal interface has socket pair {(IP\textsubscript{in},PORT\textsubscript{in},IP\textsubscript{ex},PORT\textsubscript{ex})}.

@@

\subsection{NAT classes}\label{sec:nat}
NAT is classified w.r.t. \emph{filtering rule} and \emph{allocation rule} which says what happens after receiving
packet from an external and an internal network respectively.

Common NAT categorization is according to~\citep{rfc3489}: a) Full Cone
b) IP Restricted Cone, c) Port Restricted Cone, d) Symmetric.

@@

\paragraph{Port restricted cone NAT} ~\\
Allocation table $\mathbb{AT} \subseteq \{\mathbb{IP}_{in} \times \mathbb{PORT} \times \mathbb{IP}_{ex} \times \mathbb{PORT}\} \times \mathbb{PORT}$ 
holds associations between internal hosts socket, external host socket and mapped external NAT port. 
% Rule is created once the connection from the internal host is initiated to an arbitrary external address.
Filtering rule is:
\begin{align*}
& PASS((\text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}}, \text{IP\textsubscript{nat}}, \text{PORT\textsubscript{nat}} )) \Leftrightarrow \\
&  \exists \; \text{IP\textsubscript{in}} \in \mathbb{IP}, \; \text{PORT\textsubscript{in}} \in \mathbb{PORT}:\\
& ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}, \text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}}), \text{PORT\textsubscript{nat}} ) \in \mathbb{AT}
\end{align*}

\paragraph{Symmetric NAT} ~\\
Allocation table $\mathbb{AT}$ and filtering rule is the same as in Port restricted Cone NAT. The difference lies in the way 
the new entries in the allocation table are created. For Cone NAT holds that multiple entries in the allocation table 
can have the same external port assigned, while Symmetric NAT creates new allocation entries under certain circumstances.

\emph{Allocation function (AF)} defines port allocation rules for new connections (not in $\mathbb{AT}$). Assume $\mathbb{AT}=\emptyset$, for simplicity. 
AF can be either \emph{predictable} or \emph{random}. In our setting predictable AF is assumed, e.g., \emph{incremental}. 
If previously allocated port was $\text{PORT\textsubscript{nat}}$
the next port will be $\text{PORT\textsubscript{nat}}+\Delta$, where $\Delta=1$ usually.
If the chosen port is already taken, AF iterates until the free one is found.

@@

\section{UDP hole punching with STUN}
All algorithms in our paper make use of the same principle, \emph{UDP hole punching}. 
Internet hosts cannot send packet to the internal host directly. 
Communication has to be initiated from the inside\footnote{Not taking
pre-set port forwarding, DMZ, uPnP, etc... into consideration. }, what is problematic if two nodes behind 
NAT wants to communicate.

Assume two parties, A and B with addresses IP\textsuperscript{A} and IP\textsuperscript{B} respectively,
A and B are behind public NAT with addresses $\text{IP}^{\text{A}}_{\text{nat}}$ and $\text{IP}^{\text{B}}_{\text{nat}}$
respectively. By default an internal host does not known its external address and NAT type thus technique like 
STUN\footnote{Core idea: publicly available service running on 2 IP addresses and 2 ports 
answering questions like: what is my IP address and port? Change IP address of response if I'll get it, etc...} 
is used to learn it. In case of incremental NAT it is also needed to determine $\Delta$ and the next port
that is likely to be allocated for a new connection $\text{PORT}^{\text{A}}_{\text{nat}}$ for A and $\text{PORT}^{\text{B}}_{\text{nat}}$
for B. The period between determining this information, sending to the other party and starting the algorithm is
denoted as an \emph{initialization phase} or a \emph{silent period}\ignore{\footnote{Since the algorithm itself does nothing.}} of the NAT traversal algorithm.

@@

\section{Theoretical network model}
In order to evaluate algorithms in multiple environments with different parameters (workload, number of clients, etc.)
we decided to model network behavior with stochastic process. In particular we choose a Poisson process, $\{N(t), t\geq0\}$ 
that is used in queueing theory to model arrivals of requests to the server queue~\citep{Nelson:1995:PSP:207382}.

The core idea of NAT traversal is to predict the next allocated external port by NAT, thus from this perspective 
it is important to model how NAT's internal state changes over time, i.e., how many external ports 
were allocated in a given time interval. Thus the workload is important parameter for predicting an allocated port.

We model internal network connected to NAT w.r.t. newly created connections (new port allocation) as a 
time homogeneous Poisson process $Po(\lambda)$. By setting $\lambda$ we can model 
different workload of the network and evaluate algorithms with respect to the various parameters e.g., 
probability of an establishing connection.

@@

Moreover we assume incremental NAT with $\Delta=1$ on both sides. If we sample the NAT 
state\footnote{STUN request to determine current port number, it creates a new NAT allocation record.}
each $T=10$~milliseconds the NAT state can be modeled as a random process $\{C_i | C_i \in \mathbb{N}, C_i \geq i, i\geq0\}$ where:
\[
C_i = \begin{dcases*}
         0 & if $i=0$ \\
         C_{i-1} + 1 + X_i, \; X_i \sim Po(\lambda T) & otherwise 
        \end{dcases*}
\]
Rewritten as $C_i = i + \sum_{j=1}^{i}X_j$. Expected value $E[C_i] = i + E[\sum_{j=1}^{i}X_i] = i (1+\lambda T)$.

\paragraph{Port pool exhaustion.} NAT timeout\footnote{Allocation is deleted if no packet is detected in this time interval.} is assumed to be 3~minutes, 
what is common value in practice. This constrains $\lambda$ to the interval $[0, 0.36]$, intuitively from $0$ to $360$ new connections in $1$~s on average. 
If it is higher the NAT is rendered as unusable since the port pool is exhausted\footnote{$P[X > 65535] = 0.0019, X \sim Po(180000 \cdot 0.36)$.} quickly.

\section{Generic algorithm structure}
Algorithms we propose have a common structure. In order to maximize probability of an establishing
a connection (i.e., success rate) it proceeds in discrete \emph{steps}. In each step it 
tries to open a new hole on the local NAT and/or to predict corresponding port allocated on the destination
NAT.

\begin{mydef}
Symmetric NAT traversal algorithm is a mapping 
$\mathcal{A}: \mathcal{U} \times \mathcal{S} \times \mathcal{M} \rightarrow \mathcal{P}_{src} \times \mathcal{P}_{dst}$, 
where \\
\begin{compactitem}
\item $\mathcal{U}=\{0,1\}$ is set of communicating parties.
\item $\mathcal{S} = \{0, 1, \dots, n-1\}, \; n \in \mathbb{N}$ is a step of the algorithm.
\item $\mathcal{M}$ is set of vectors of parameters of the environment (i.e., network model 
parameters and additional information obtained in initial phase). 
In our setting $\mathcal{M} = \{ (\lambda, t_s) | \lambda~>~0, t\geq~0 \}$. Value $t_s$ is length of the silent period. %For simplicity $\mathcal{M} = \{ \lambda | \lambda > 0 \}$.
\item $\mathcal{P}_{src} \subseteq \mathbb{N}$ and $\mathcal{P}_{dst} \subseteq \mathbb{N}$
are sets of source and destination ports respectively.
\end{compactitem}
\end{mydef}
If the third parameter is not important for the algorithm it can be ignored in the notation.

Prior algorithm run it is needed to 
determine NAT state (for UDP hole punching), this is assumed to be done by some technique like STUN. For simplicity we assume the port numbers
starts from $0$. The length of the silent period is also known to the algorithm prior execution in order to estimate the current state of 
the NAT, e.g., difference between the last measurement and just before algorithm execution.

\paragraph{Intuition.} 
Line $A(0,1) = (2,3)$ means party $0$ in step $1$ sent packet from source port $2$ to destination port $3$
of the party $1$.

\paragraph{Time.} 
%The algorithm proceeds in discrete time steps. 
Algorithm starts in time $t_0 = 0$~ms and each step $i \in \mathcal{S}$
happens in time $t_i = iT$ for some $T$. If it is not mentioned otherwise, $T = 10$~ms.

@@

\paragraph{Algorithm log \& successful run.} IP addresses of the hosts and NAT are known and fixed prior algorithm run thus omitted
from the notation for simplicity. For the following text is assumed maximum number of steps of the algorithm $s_{max}=1000$.

@@

\begin{compactitem}
 \item $(p_1, p_2) = \mathcal{A}(0, s, (\lambda, t_s)) $ are generated by the algorithm $\mathcal{A}$ for some 
       parameters $(\lambda, t_s)$ in some step $s$ of the algorithm $\mathcal{A}$.
 \item $p_3$ is an allocated port by $NAT^A$ for request $(p_1, p_2)$ in a given step $s$. In our network model
       $p_3=C_s=C_{s-1} + 1 + X, X~\sim~Po(\lambda T)$.
 \item Algorithm log for host $B$ is defined analogically.
 \item Furthermore from our setting the following formula holds:\footnote{From definition of a port sensitive symmetric NAT.}
\end{compactitem} %
%
% \begin{compactitem}
% \item $p_1$ is a source port.
% \item $p_2$ is a destination port.
% \item $p_3$ is an allocated port by local NAT for this request.
% \item following formula holds\footnote{from definition of a port sensitive symmetric NAT}
% \end{compactitem}
%
\begin{align*}
& (p_1, p_2, p_3) \in \mathcal{L}^A_{\mathcal{A}}: \forall (p_1^{\prime}, p_2^{\prime}, p_3^{\prime}) \in \mathcal{L}^{A}_{\mathcal{A}} :\\
& p_3 = p_3^{\prime} \Leftrightarrow \left( p_1=p^{\prime}_1 \wedge p_2=p^{\prime}_2 \right)
\end{align*}
\end{mydef}

\begin{mydef}
\label{def:match}
Algorithm $\mathcal{A}$ run is successful, i.e., established a direct connection between $A,B$ if\footnote{Derived from UDP hole punching.}:
\begin{align*}
& MATCH(\mathcal{L}^{A}_{\mathcal{A}}, \mathcal{L}^{B}_{\mathcal{A}}) \Leftrightarrow \exists(p_1, p_2, p_3) \in \mathcal{L}^{A}_{\mathcal{A}}: \\
& \exists(p_1^{\prime}, p_2^{\prime}, p_3^{\prime}) \in \mathcal{L}^{B}_{\mathcal{A}}: p_2 = p_3^{\prime} \wedge p_2^{\prime} = p_3
%
%  \in \mathcal{L}^A_{\mathcal{A}}: \forall (p_1^{\prime}, p_2^{\prime}, p_3^{\prime}) \in \mathcal{L}^{A}_{\mathcal{A}} :\\
%& p_3 = p_3^{\prime} \left( \Leftrightarrow p_1=p^{\prime}_1 \wedge p_2=p^{\prime}_2 \right)
\end{align*}
\end{mydef}
The task is to find an algorithm $\mathcal{A}$ such that the probability the $MATCH(\mathcal{L}^{A}_{\mathcal{A}}, \mathcal{L}^{B}_{\mathcal{A}})$
holds is maximized for minimal number of steps needed.

\subsection{Algorithm I -- baby-step, giant-step}
We propose the algorithm with high success rate for small network loads, i.e.,
$\lambda \leq 0.03, T=10$~ms. When error becomes larger,
algorithm step is not fast enough to compensate errors and thus fails for larger $\lambda$.

@@

Intuitively the source port is kept constant, one side of the protocol scans ports sequentially, another
side of the protocol scans ports with step 2.

@@

To validate our results we implemented Android application using baby-step giant-step algorithm
to traverse symmetric carrier-grade NAT the mobile phone operator was using. The implementation
works according to expectations.

>>>>

\urldef{\mailsa}\path|xklinec@fi.muni.cz|
\urldef{\mailsb}\path|matyas@fi.muni.cz|
\author{Du\v{s}an Klinec \and Vashek Matyáš}
\institute{Faculty of Informatics, Masaryk University, Brno, Czech Republic\\
\email{\{xklinec,matyas\}@fi.muni.cz}}

@@

\begin{abstract}
\noindent Network Address Translators cause troubles for VoIP and other P2P services since central servers are needed for communication.
Presence of such (potentially malicious) hosts in a communication path is not desired, mainly due to increased cost, poor link quality 
and security consequences.
Several solutions exist for traversing NAT, but symmetric case is still problematic. We propose algorithms using single source port for 
symmetric NAT traversal, each with different properties and applicability. 
Our paper investigates the NAT traversal problem by stating the problem formally and modeling a network connected to 
NAT as a random process, in order to evaluate algorithms in different settings.
\end{abstract}

@@

\section{Introduction}
Network Address Translators (abbreviated as NATs) have been increasing in popularity in recent years, 
mainly as a mean of slowing down IPv4 address space depletion. On one hand, NAT offers
the ability to share a public, globally routable IP address among many internal network hosts. On 
the other hand, it breaks the basic end-to-end principle of the Internet and thus favors server-client 
communication model for clients connected behind NAT.

In many applications, a peer-to-peer (P2P) connection is of crucial importance and the quality of a service depends on 
its parameters heavily. Main motivation for this work is Voice-over-IP (VoIP) where direct connection 
between communicating parties is necessary. If P2P connection is not possible, relay 
servers have to be used (TURN~\citep{rfc5766} protocol is used). The presence 
of another node in the communication path can have negative effects on the communication channel quality 
(increased latency, jitter, lower bandwidth), the relay server is a potential point of failure and 
has to scale with increasing demands of the network, etc.
Thus relay servers require non-negligible resources and become expensive infrastructure components for VoIP and other service providers where establishing a direct connection between parties is of high importance.

From the attacker's perspective, relay servers can be seen as potential allies when trying to sniff and log all 
communication. NAT (especially symmetric one, used as a carrier-grade border firewall in large networks)
forces communicating parties to use such central servers
for communication, making monitoring easier for an attacker. Traversal algorithms have to collaborate 
with central servers in such a way they create a direct link, not using servers any further. 
% 
% some way. They use it to create a direct link, not using servers any further.

In this paper, we focused on a symmetric NAT with a predictable port allocation function. We propose 
new algorithms for establishing direct connection between hosts behind such NATs that are effective and 
have low system requirements.

@@

Some devices with NAT functionality provide special interface enabling traversal, for example NAT Port Mapping Protocol~\citep{rfc6886}. Yet in general, it is not widely deployed and it cannot be relied upon in a general case.

ICE\ignore{~\citep{rfc5245}} is used as a technique to traverse NAT in most applications nowadays, 
although it does not work with symmetric NATs. ICE mainly uses STUN~\citep{rfc5389} and TURN\ignore{~\citep{rfc5766}} for its operation.

% \section{Related work}
Some interesting algorithms for traversing symmetric NAT were proposed by Y.~Takeda in \citep{takeda}, while the case where both clients are behind symmetric NAT
with a port sensitive allocation function is not covered. Takeda mentions that sending more packets does not increase a probability of establishing a new connection.
Our motivation is to show that it is actually possible.

RFC~5128 \citep{rfc5128} maps current state of the art in NAT traversal, but does not cover our case. 
The research by Wei~\emph{et~al.} in \citep{wei}., is devoted to symmetric NAT traversal, but their algorithm is using another technique compared to ours, and the the downside of their algorithm comes with the Time To Live~(TTL) parameter modification of the IP~packet. This is usually considered
as a privileged operation in the operating system and super-user privileges are required for this modification.

Wang~\emph{et~al.} \citep{Wang:2006:RSN:1156422.1156550} propose a NAT traversal algorithm for the scenario where two hosts are
behind symmetric NATs, based on a changing of a source port, while we explore options with a constant source port.

In our work we study the situation where two parties -- $A, B$ -- try to establish a direct connection, both of them are behind 
a symmetric NAT (possibly a cascade of such NATs) with a port sensitive, predictable allocation function. Other scenarios with
predictable allocation function are quite easy to solve and covered in previous research.

\section{Network Address Translation}
The notation used for the rest of our paper follows.

@@

\begin{mydef}
% \begin{align*}
$\mathbb{SC} = \mathbb{IP} \times \mathbb{PORT}$
% \end{align*}%
 is a Socket. $\mathbb{IP}$ is the set of all possible IP addresses and $\mathbb{PORT}$ 
is the set of all possible ports, basically $\mathbb{PORT} = \{1,2,\dots,65535\}$.
% \begin{align*}
$\mathbb{SP} = \mathbb{IP} \times \mathbb{PORT} \times \mathbb{IP} \times \mathbb{PORT}$
% \end{align*}%
 is a Socket Pair.
\end{mydef}
% 
% \begin{mydef}
% \begin{align*}
% \mathbb{AT} \subseteq \mathbb{SP} \times \mathbb{PORT}
% \end{align*} is an allocation table of the NAT. Second component is external port allocated for particular socket pair.
% \end{mydef}
% Allocation table is a set of all socket pairs and corresponding allocated external port.

\paragraph{NAT and communication.}
NAT connects two networks, denoted as {\em internal} and {\em external}, connected to an internal and an external NAT interface, respectively.
Hosts have addresses $\text{IP\textsubscript{in}} \in \mathbb{IP}_{in} \subseteq \mathbb{IP}$ 
and $\text{IP\textsubscript{ex}} \in \mathbb{IP}_{ex} \subseteq \mathbb{IP}$ in the internal and the external network,
respectively, where $\mathbb{IP}_{in} \cap \mathbb{IP}_{ex} = \emptyset$.
If it is not stated otherwise the external NAT address\footnote{We assume NAT has only one external address for simplicity.} is denoted as 
$\text{IP\textsubscript{nat}} \in \mathbb{IP}_{ex}$.

Packet received on the external interface has the socket pair {(IP\textsubscript{ex},PORT\textsubscript{ex},IP\textsubscript{nat},PORT\textsubscript{nat})}.
Packet received on the internal interface has the socket pair {(IP\textsubscript{in},PORT\textsubscript{in},IP\textsubscript{ex},PORT\textsubscript{ex})}.

@@

\subsection{NAT classes}\label{sec:nat}
NAT is classified w.r.t. the \emph{filtering rule} and the \emph{allocation rule}, which state what happens after receiving
packet from an external and an internal network, respectively.

Common NAT categorization is according to~\citep{rfc3489}: a) Full Cone;
b) IP Restricted Cone; c) Port Restricted Cone; d) Symmetric.

@@

\paragraph{Port Restricted Cone NAT} ~\\
Allocation table $\mathbb{AT} \subseteq \{\mathbb{IP}_{in} \times \mathbb{PORT} \times \mathbb{IP}_{ex} \times \mathbb{PORT}\} \times \mathbb{PORT}$ 
holds associations between the internal host socket, external host socket and mapped external NAT port. 
% Rule is created once the connection from the internal host is initiated to an arbitrary external address.
The filtering rule is:
\begin{align*}
& PASS((\text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}}, \text{IP\textsubscript{nat}}, \text{PORT\textsubscript{nat}} )) \Leftrightarrow \\
&  \exists \; \text{IP\textsubscript{in}} \in \mathbb{IP}, \; \text{PORT\textsubscript{in}} \in \mathbb{PORT}:\\
& ((\text{IP\textsubscript{in}}, \text{PORT\textsubscript{in}}, \text{IP\textsubscript{ex}}, \text{PORT\textsubscript{ex}}), \text{PORT\textsubscript{nat}} ) \in \mathbb{AT}
\end{align*}

\paragraph{Symmetric NAT} ~\\
Allocation table $\mathbb{AT}$ and filtering rule are the same as in the Port Restricted Cone NAT. The difference lies in the way 
the new entries in the allocation table are created. For the Cone NAT holds that multiple entries in the allocation table 
can have the same external port assigned, while the Symmetric NAT creates new allocation entries under certain circumstances.

\emph{Allocation function (AF)} defines port allocation rules for new connections (not in $\mathbb{AT}$). Assume $\mathbb{AT}=\emptyset$, for simplicity. 
AF can be either \emph{predictable} or \emph{random}. In our setting, a predictable AF is assumed, e.g., \emph{incremental}. 
If the previously allocated port was $\text{PORT\textsubscript{nat}}$ then
the next port will be $\text{PORT\textsubscript{nat}}+\Delta$, where typically $\Delta=1$.
If the chosen port is already taken, AF iterates until a free one is found.

@@

\section{UDP hole punching with STUN}
All algorithms in our paper make use of the same principle, \emph{UDP hole punching}. 
Internet hosts cannot send packets to the internal host directly. 
Communication has to be initiated from the inside\footnote{Not taking
pre-set port forwarding, DMZ, uPnP, etc. into consideration. }, which is problematic if two nodes behind 
NAT want to communicate.

Assume two parties, A and B with addresses IP\textsuperscript{A} and IP\textsuperscript{B}, respectively.
A and B are behind a public NAT with addresses $\text{IP}^{\text{A}}_{\text{nat}}$ and $\text{IP}^{\text{B}}_{\text{nat}}$,
respectively. By default an internal host does not know its external address and NAT type, thus a technique like 
STUN\footnote{Core idea: publicly available service running on 2 IP addresses and 2 ports 
answering questions like: what is my IP address and port? Change IP address of response if Il get it, etc.} 
is used to learn it. In case of incremental NAT it is also necessary to determine $\Delta$ and the next port
that is likely to be allocated for a new connection $\text{PORT}^{\text{A}}_{\text{nat}}$ for A and $\text{PORT}^{\text{B}}_{\text{nat}}$
for B. The period between determining this information, sending to the other party and starting the algorithm is
denoted as an \emph{initialization phase} or a \emph{silent period}\ignore{\footnote{Since the algorithm itself does nothing.}} of the NAT traversal algorithm.

@@

\section{Theoretical network model}
In order to evaluate algorithms in multiple environments with different parameters (workload, number of clients, etc.)
we decided to model network behavior with a stochastic process. In particular, we chose a Poisson process, $\{N(t), t\geq0\}$ 
that is used in queueing theory to model arrivals of requests to the server queue~\citep{Nelson:1995:PSP:207382}.

The core idea of NAT traversal is to predict the next allocated external port by NAT, thus from this perspective 
it is important to model how the NAT internal state changes over time, i.e., how many external ports 
have been allocated in a given time interval. Thus the workload is an important parameter for predicting the allocated port.

We model the internal network connected to NAT w.r.t. newly created connections (new port allocation) as a 
time homogeneous Poisson process $Po(\lambda)$. By setting $\lambda$ we can model 
different workloads of the network and evaluate algorithms w.r.t. various parameters, e.g., 
probability of establishing a connection.

@@

Moreover, we assume an incremental NAT with $\Delta=1$ on both sides. If we sample the NAT 
state\footnote{STUN requests to determine current port number, it creates a new NAT allocation record.}
each $T=10$~milliseconds the NAT state can be modeled as a random process $\{C_i | C_i \in \mathbb{N}, C_i \geq i, i\geq0\}$ where:
\[
C_i = \begin{dcases*}
         0 & if $i=0$ \\
         C_{i-1} + 1 + X_i, \; X_i \sim Po(\lambda T) & otherwise 
        \end{dcases*}
\]
Rewritten as $C_i = i + \sum_{j=1}^{i}X_j$. Expected value $E[C_i] = i + E[\sum_{j=1}^{i}X_i] = i (1+\lambda T)$.

\paragraph{Port pool exhaustion.} NAT timeout\footnote{Allocation is deleted if no packet is detected within this time interval.} is assumed to be 3~minutes, 
which is a typical setting in practice. This constrains $\lambda$ to the interval $[0, 0.36]$, intuitively from $0$ to $360$ new connections in $1$~s on average. 
If it is higher then NAT is rendered as unusable since the port pool is exhausted\footnote{$P[X > 65535] = 0.0019, X \sim Po(180000 \cdot 0.36)$.} quickly.

\section{Generic algorithm structure}
Algorithms that we propose have a common structure. In order to maximize the probability of establishing
a connection (i.e., success rate) it proceeds in discrete \emph{steps}. In each step it 
tries to open a new hole in the local NAT and/or to predict the corresponding port allocated in the destination
NAT.

\begin{mydef}
Symmetric NAT traversal algorithm is a mapping 
$\mathcal{A}: \mathcal{U} \times \mathcal{S} \times \mathcal{M} \rightarrow \mathcal{P}_{src} \times \mathcal{P}_{dst}$, 
where \\
\begin{compactitem}
\item $\mathcal{U}=\{0,1\}$ is the set of communicating parties.
\item $\mathcal{S} = \{0, 1, \dots, n-1\}, \; n \in \mathbb{N}$ is a step of the algorithm.
\item $\mathcal{M}$ is the set of vectors of parameters of the environment (i.e., network model 
parameters and additional information obtained in initial phase). 
In our setting $\mathcal{M} = \{ (\lambda, t_s) | \lambda~>~0, t\geq~0 \}$. Value $t_s$ denotes the length of the silent period. %For simplicity $\mathcal{M} = \{ \lambda | \lambda > 0 \}$.
\item $\mathcal{P}_{src} \subseteq \mathbb{N}$ and $\mathcal{P}_{dst} \subseteq \mathbb{N}$
are sets of source and destination ports respectively.
\end{compactitem}
\end{mydef}
In case $\mathcal{M}$ is not not important for the algorithm it can be ignored in the notation.
%If the third parameter is not important for the algorithm it can be ignored in the notation.

Prior algorithm run it is needed to 
determine the NAT state (for UDP hole punching), this is assumed to be done by some technique like STUN. For simplicity, we assume that port numbers
start from $0$. The length of the silent period is also known to the algorithm prior execution in order to estimate the current state of 
the NAT, e.g., difference between the last measurement and just before actual algorithm execution.

\paragraph{Notation.}
Line $A(0,1) = (2,3)$ means party $0$ in step $1$ sent packet from source port $2$ to destination port $3$
of the party $1$.

\paragraph{Time.} 
%The algorithm proceeds in discrete time steps. 
Algorithm starts in time $t_0 = 0$~ms and each step $i \in \mathcal{S}$
happens in time $t_i = iT$ for some $T$. If not mentioned otherwise, $T = 10$~ms.

@@

\paragraph{Algorithm log \& successful run.} IP addresses of the hosts and NAT are known and fixed prior to the algorithm run, and are thus omitted from the notation for simplicity. For the following text we assume the maximum number of steps of the algorithm $s_{max}=1000$.

@@

\begin{compactitem}
 \item $(p_1, p_2) = \mathcal{A}(0, s, (\lambda, t_s)) $ are generated by the algorithm $\mathcal{A}$ for some 
       parameters $(\lambda, t_s)$ in some step $s$ of the algorithm $\mathcal{A}$;
 \item $p_3$ is an allocated port by $NAT^A$ for request $(p_1, p_2)$ in a given step $s$. In our network model
       $p_3=C_s=C_{s-1} + 1 + X, X~\sim~Po(\lambda T)$;
 \item algorithm log for host $B$ is defined analogically;
 \item furthermore from our setting the following formula holds:\footnote{From definition of a port sensitive symmetric NAT.}
\end{compactitem} %
%
% \begin{compactitem}
% \item $p_1$ is a source port.
% \item $p_2$ is a destination port.
% \item $p_3$ is an allocated port by local NAT for this request.
% \item following formula holds\footnote{from definition of a port sensitive symmetric NAT}
% \end{compactitem}
%
\begin{align*}
& (p_1, p_2, p_3) \in \mathcal{L}^A_{\mathcal{A}}: \forall (p_1^{\prime}, p_2^{\prime}, p_3^{\prime}) \in \mathcal{L}^{A}_{\mathcal{A}} :\\
& p_3 = p_3^{\prime} \Leftrightarrow \left( p_1=p^{\prime}_1 \wedge p_2=p^{\prime}_2 \right)
\end{align*}
\end{mydef}

\begin{mydef}
\label{def:match}
Algorithm $\mathcal{A}$ run is successful, i.e., established a direct connection between $A,B$ if~\footnote{Derived from UDP hole punching.}:
\begin{align*}
& MATCH(\mathcal{L}^{A}_{\mathcal{A}}, \mathcal{L}^{B}_{\mathcal{A}}) \Leftrightarrow \exists(p_1, p_2, p_3) \in \mathcal{L}^{A}_{\mathcal{A}}: \\
& \exists(p_1^{\prime}, p_2^{\prime}, p_3^{\prime}) \in \mathcal{L}^{B}_{\mathcal{A}}: p_2 = p_3^{\prime} \wedge p_2^{\prime} = p_3.
%
%  \in \mathcal{L}^A_{\mathcal{A}}: \forall (p_1^{\prime}, p_2^{\prime}, p_3^{\prime}) \in \mathcal{L}^{A}_{\mathcal{A}} :\\
%& p_3 = p_3^{\prime} \left( \Leftrightarrow p_1=p^{\prime}_1 \wedge p_2=p^{\prime}_2 \right)
\end{align*}
\end{mydef}
The task is to find an algorithm $\mathcal{A}$ such that the probability $MATCH(\mathcal{L}^{A}_{\mathcal{A}}, \mathcal{L}^{B}_{\mathcal{A}})$
holds is maximized for minimal number of steps needed.

\subsection{Algorithm I -- baby-step, giant-step}
We propose the algorithm with high success rate for small network loads, i.e.,
$\lambda \leq 0.03, T=10$~ms. When the error becomes larger, the
algorithm step is not fast enough to compensate errors and thus fails for a larger $\lambda$.

@@

Intuitively, the source port is kept constant, one side of the protocol scans ports sequentially, another
side of the protocol scans ports with step 2.

@@

To validate our results, we implemented Android application using our baby-step giant-step algorithm
to traverse symmetric carrier-grade NAT that our mobile phone operator was using. The implementation
works according to expectations.
